Supplement to 'Delayed effects of climate on vital rates leads to demographic divergence in Amazonian forest fragments

### 2021-05-13 doc/supplemental.Rmd
finalize supplemental doc

withr::with_dir(here::here(), {
  targets::tar_load(c(
    normals_plot,
    embrapa_wide,
    xa_spei,
    trmm_spei,
    gspei
    ))
})

library(tidyverse)
library(patchwork)
library(flextable)
library(tsibble)
library(lubridate)
library(latex2exp)
library(colorspace)
library(glue)
source(here::here("R", "utils.R"))
```

@@

Map caption here

@@

Figure S1.
Weather normals for Manaus, Brazil (3º6'S, 60º1'W) for 1981--2010.
Precipitation (a) shows a marked dry seasons from June through October.
Temperature (b) varies less throughout the year.
Mean monthly temperature is shown in the solid red line and monthly minimum and maximum temperatures are shown with the lower and upper dashed lines, respectively.
Data from Brazilian National Institute of Meterology (INMET).
<!--# citation needed? -->

@@

Figure S2.
Three-month SPEI at BDFFP calculated using 4 datasets.
"EMBRAPA" (red) is the average of SPEI calculated using precipitation and evapotranspiration from two weather stations near BDFFP---Estacao Rio Preto da Eva ( -2.65º, -59.65º) and Estacao Manaus (-2.65º, -60.35º).
The data was accessed from the EMBRAPA Infoclima data portal (<https://www.cnpaf.embrapa.br/infoclima/>).
"Global SPEI" (green) is from the SPEI global database[@vicente-serranoNewGlobalGridded2010; @begueriaMultiscalarGlobalDrought2010].
"TRMM" (blue) SPEI uses precipitation data from TRMM [@trmmTMPA3B432011] and evapotranspiration data from @xavierDailyGriddedMeteorological2016 to calculate 3-month SPEI.
Note that TRMM data only goes back to 1998.
@xavierDailyGriddedMeteorological2016 is the SPEI source described in the main text.
Drought categories mild, moderate, severe, and extreme drought are represented by yellow, goldenrod, orange, and red panels, respectively.

@@

>>>>

withr::with_dir(here::here(), {
  targets::tar_load(c(
    normals_plot,
    embrapa_wide,
    xa_spei,
    trmm_spei,
    gspei,
    la_data
    ))
})

library(tidyverse)
library(patchwork)
library(flextable)
library(tsibble)
library(lubridate)
library(latex2exp)
library(colorspace)
library(glue)
library(performance)
source(here::here("R", "utils.R"))
```

@@

Figure S1.
Map caption here

@@

Figure S2.
Weather normals for Manaus, Brazil (3º6'S, 60º1'W) for 1981--2010.
Precipitation (a) shows a marked dry seasons from June through October.
Temperature (b) varies less throughout the year.
Mean monthly temperature is shown in the solid red line and monthly minimum and maximum temperatures are shown with the lower and upper dashed lines, respectively.
Data from Brazilian National Institute of Meterology (INMET).
<!--# citation needed? -->

@@

Figure S3.
Three-month SPEI at BDFFP calculated using 4 datasets.
"EMBRAPA" (red) is the average of SPEI calculated using precipitation and evapotranspiration from two weather stations near BDFFP---Estacao Rio Preto da Eva ( -2.65º, -59.65º) and Estacao Manaus (-2.65º, -60.35º).
The data was accessed from the EMBRAPA Infoclima data portal (<https://www.cnpaf.embrapa.br/infoclima/>).
"Global SPEI" (green) is from the SPEI global database[@vicente-serranoNewGlobalGridded2010; @begueriaMultiscalarGlobalDrought2010].
"TRMM" (blue) SPEI uses precipitation data from TRMM [@trmmTMPA3B432011] and evapotranspiration data from @xavierDailyGriddedMeteorological2016 to calculate 3-month SPEI.
Note that TRMM data only goes back to 1998.
@xavierDailyGriddedMeteorological2016 is the SPEI source described in the main text.
Drought categories mild, moderate, severe, and extreme drought are represented by yellow, goldenrod, orange, and red panels, respectively.

@@

```{r}
m_ht <-   lm(log_leaf_area ~ log_ht,   data = la_data)
m_shts <- lm(log_leaf_area ~ log_shts, data = la_data)
m_size <- lm(log_leaf_area ~ log_size, data = la_data)
```

```{r include=FALSE}
check_model(m_ht)
check_model(m_shts)
check_model(m_size)
```

Table S2.
Using data from transplant experiments of *Heliconia acuminata* [@brunaExperimentalAssessmentHeliconia2002 <!--# and doi for unpublished data -->] we fit linear regressions with log(leaf area) as the response variable and either log(height), log(shoot number) or log(shoot number x height) as the predictor.
Using the log of size (shoot number $\times$ height) as a predictor gives the lowest AIC, BIC, root mean squared error (RMSE) and highest $R^2$ indicating that it is the best proxy for predicting log(leaf area) as a response.

```{r size-table}
compare_performance(m_ht, m_shts, m_size, metrics = "common", rank = TRUE) %>%
  select(-Model) %>% 
  add_column(
    "Model" = c(
      "log(leaf area) ~ log(shoot number x height)",
      "log(leaf area) ~ log(height)",
      "log(leaf area) ~ log(shoot number)"
    )
  ) %>%
  mutate(dAIC = AIC - min(AIC), dBIC = BIC- min(BIC)) %>% 
  mutate(across(c(dAIC, dBIC, R2), ~round(.x, 2)), RMSE = round(RMSE, 3)) %>% 
  select(Model, dAIC, dBIC, R2, RMSE) %>% 
  flextable() %>% 
  set_table_properties(layout = "autofit") 
```
